---
title: "House Selling Prices"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    df_print: paged
---

# {.tabset .tabset-pills}


```{r message=FALSE, warning=FALSE}
library(tidyverse)
# library(car)
library(ggpubr)
library(pander)
library(DT)
```

## Analysis

### Introduction

This analysis accurately predicts the house "SalePrice" base on a Kaggle dataset based on a regression model.

For this, I created made some data cleaning and fitted a regression into it.

__The Adjusted R^2 is 0.8219 with an R^2 of 0.809.__


```{r message=FALSE, warning=FALSE}

dat <- read.csv("train.csv")

# datatable(dat, options=list(lengthMenu = c(10)), extensions="Buttons")
# 
# datatable(head(dat, 20), options = list(
#   columnDefs = list(list(className = 'dt-center', targets = 5)),
#   pageLength = 5,
#   lengthMenu = c(5, 10, 15, 20)))


set.seed(121)

num_rows <- 1000 #1460 total
keep <- sample(1:nrow(dat), num_rows)

dat2 <- dat %>% 
  mutate(
    TotalSF = `X1stFlrSF` + `X2ndFlrSF` + `TotalBsmtSF`,
    OverallQual = case_when(
           OverallQual >= 9 ~ "Qual9Up",
           OverallQual <= 4 ~ "Qual4Less",
           TRUE ~ as.character(paste0("Qual",OverallQual))),
    SalePriceLog = log(SalePrice),
    MSZoning2 = case_when(
        MSZoning %in% c("RH", "RL", "RM", "RP") ~ "ZoningR",
        MSZoning == "C (all)" ~ "ZoningC",
        TRUE ~paste0("Zoning", MSZoning)),
    GarageCars = case_when(
        GarageCars >= 3 ~ "Garage3Up",
        TRUE ~ as.character(paste0("Garage", GarageCars))),
    TotalSF = ifelse(TotalSF > 7000, base::mean(TotalSF), TotalSF),
    OverallQual = as.factor(OverallQual),
    GarageCars = as.factor(GarageCars),
    MSZoning2 = as.factor(MSZoning2)) %>% 
  select(TotalSF, SalePriceLog,SalePrice, MSZoning2, GarageCars, OverallQual) %>% 
  mutate(
    rownum = 1:n(), value = 1) %>% 
    spread(OverallQual, value, fill = 0) %>% 
    mutate(value = 1) %>% 
    spread(MSZoning2, value, fill = 0) %>% 
    mutate(value = 1) %>% 
    spread(GarageCars, value, fill = 0)

mytrain <- dat2[keep, ] #Use this in the lm(..., data=mytrain)

mytest <- dat2[-keep, ] #Use this in the predict(..., newdata=mytest)

lm1 <- lm(SalePriceLog~
TotalSF
+Qual4Less
+Qual5
+Qual6 
+Qual7 
+Qual8 
+Qual4Less:Garage2:TotalSF
+Qual5:Garage1:TotalSF 
+Qual5:Garage2 
+Qual6:Garage2 
+Qual8:Garage3Up:TotalSF
,data = mytrain)

yht <- predict(lm1, newdata = mytest)

# Compute y-bar
ybar <- mean(mytest$SalePriceLog) #Yi is given by Ynew from the new sample of data
  
# Compute SSTO
SSTO <- sum( (mytest$SalePriceLog - ybar)^2 )
  
# Compute SSE for each model using y - yhat
SSEt <- sum( (mytest$SalePriceLog - yht)^2 )

  # Compute R-squared for each
  rst <- 1 - SSEt/SSTO
  
# Compute adjusted R-squared for each
n <- length(mytest$SalePriceLog) #sample size
pt <- length(coef(lm1)) #num. parameters in model
rsta <- 1 - (n-1)/(n-pt)*SSEt/SSTO

my_output_table <- data.frame(
    `Original R2`           = summary(lm1)$r.squared,
    `Orig. Adj. R-squared` = summary(lm1)$adj.r.squared,
    `Validation R-squared` = rst,
    `Validation Adj. R^2`  = rsta)

colnames(my_output_table) <- c("Original $R^2$", "Original Adj. $R^2$", "Validation $R^2$", "Validation Adj. $R^2$")

b <- coef(lm1)

dat3 <- dat2 %>% 
  pivot_longer(cols = starts_with("Qual"), names_to = "OverallQual") %>% 
  filter(value == 1) %>% 
  select(-value) %>% 
  pivot_longer(cols = starts_with("Garage"), names_to = "GarageCars") %>% 
  filter(value == 1)%>% 
  select(-value) %>% 
  pivot_longer(cols = starts_with("Zoning"), names_to = "MSZoning") %>% 
  filter(value == 1) %>% 
  select(-value)

dat3 %>% 
  mutate(GarageCars = case_when(
    GarageCars %in% c("Garage0", "Garage1", "Garage3Up") ~ "Garage013",
    TRUE ~ GarageCars
  )) %>% 
  ggplot(aes(TotalSF, SalePriceLog))+
  geom_point()

```

### Interpretation

Per each increase unit in square footage, the sale price increases by 0.0232% as an overall average.
The change in square footage is related to the slope value, which is 0.000232.

With the help of a graphing system, we are able to identify patterns in the data. Each of the graphs represent different situations based on the house rating and garage status.

The houses with a quality rating of 4 and 2 car garage, have a percentage change of 0.00774 per unit of total square footage which puts the total percent change at 0.031% per unit change in total square footage.

The houses with a quality rating of 5 and 1 car garage, have a percentage change of 0.029% per unit of total square footage.

The houses with a quality rating of 8 and 3 car garage, have a percentage change of 0.027% per unit of total square footage.

<br><br><br><br><br><br><br><br><br><br><br><br><br>

## Model Summary

### R-Squared Comparisson

```{r message=FALSE, warning=FALSE}

pander(my_output_table)


```

### Regression Summary

```{r message=FALSE, warning=FALSE}

pander(summary(lm1))

```

## graphing the model 

```{r message=FALSE, warning=FALSE}



p1 <- dat3 %>% 
  mutate(GarageCars = case_when(
    GarageCars %in% c("Garage0", "Garage1", "Garage3Up") ~ "Garage013",
    TRUE ~ GarageCars
  )) %>% 
  ggplot(aes(TotalSF, SalePriceLog, color = GarageCars))+
  geom_point()+
  stat_function(fun = function(x) (b[1] + b[3]) + b[2] * x + b[10] * x, color = "skyblue")+
  stat_function(fun = function(x) (b[1] + b[3]) + b[2] * x, color = "red")+
  labs(title = "Quality Rating 4 or Less")

p2 <- dat3%>% 
  mutate(GarageCars = case_when(
    GarageCars %in% c("Garage0",  "Garage3Up") ~ "Garage03",
    TRUE ~ GarageCars
  )) %>% 
  ggplot(aes(TotalSF, SalePriceLog, color = GarageCars))+
  geom_point()+
  stat_function(fun = function(x) (b[1] + b[4]) + b[2] * x + b[11] * x, color = "green")+
  stat_function(fun = function(x) (b[1] + b[4]) + b[2] * x, color = "firebrick")+
  stat_function(fun = function(x) (b[1] + b[4] + b[8]) + b[2] * x, color = "skyblue")+
  labs(title = "Quality Rating 5")
  
p3 <- dat3 %>% 
  mutate(GarageCars = case_when(
    GarageCars %in% c("Garage0", "Garage1",  "Garage3Up") ~ "Garage013",
    TRUE ~ GarageCars
  )) %>% 
  ggplot(aes(TotalSF, SalePriceLog, color = GarageCars))+
  geom_point()+
  stat_function(fun = function(x) (b[1] + b[5]) + b[2] * x + b[9], color = "skyblue")+
  stat_function(fun = function(x) (b[1] + b[5]) + b[2] * x, color = "firebrick")+
  labs(title = "Quality Rating 6")

 p4 <- dat3 %>% 
  mutate(GarageCars = case_when(
    GarageCars %in% c("Garage0", "Garage1","Garage2",  "Garage3Up") ~ "AllGarageTypes",
    TRUE ~ GarageCars
  )) %>%  
  ggplot(aes(TotalSF, SalePriceLog, color = GarageCars))+
  geom_point()+
  stat_function(fun = function(x) (b[1] + b[6]) + b[2] * x, color = "firebrick")+
  labs(title = "Quality Rating 7")

p5 <- dat3 %>% 
  mutate(GarageCars = case_when(
    GarageCars %in% c("Garage0", "Garage1","Garage2") ~ "Garage012",
    TRUE ~ GarageCars
  )) %>% 
  ggplot(aes(TotalSF, SalePriceLog, color = GarageCars))+
  geom_point()+
  stat_function(fun = function(x) (b[1] + b[7]) + b[2] * x, color = "firebrick")+
  stat_function(fun = function(x) (b[1] + b[7]) + b[2] * x + b[12] * x, color = "skyblue")+
  labs(title = "Quality Rating 8")

p1
p2
p3
p4
p5

# dat3 %>% 
#   ggplot(aes(TotalSF, SalePriceLog, color = GarageCars))+
#   geom_point()+
#   labs(title = "Quality Rating 4 or Less")+
#   facet_wrap(~interaction(
#      OverallQual
#     ,GarageCars
#     ,MSZoning
#   ))#, labeller="label_both"))

```

## Data

### Raw Data

```{r message=FALSE, warning=FALSE}

dat

```

### Clean Data

```{r message=FALSE, warning=FALSE}

dat2

```