---
title: "Predicting the Weather"
output:
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r message=FALSE, warning=FALSE}
# load packages

library(pander)
library(tidyverse)
library(mosaic)
library(car)
library(DT)

# load data
weatherData <- read.csv(
  "../../Data/WeatherPrediction.csv"
  , header = TRUE) 

# view data
# View(weatherData)
```


# {.tabset .tabset-fade}

## Introduction

### Prediction and Graph

My predicted highest  temperature in Fahrenheit for this Monday September 20th, 2021, is

<br /><br />

<center>

<span style="font-family:segoe;font-size:4em;">
82.33 &deg; F
</span>

</center>

<br /><br />

I collected weather data from the 10 days previous to this upcoming Monday, September 20th. The collected data points comes from  [www.timeanddate.com](https://www.timeanddate.com) and I extracted the day, time, temperature, weather, wind, humidity, barometer, and visibility.

I transformed the database columns and rows so the humidity data is aligned with the temperature data from two days in advance.
In other words, each row contains the temperature of the day and the humidity of the two previous days. This is done to use the humidity as a __explanatory__ variable for the Monday temperature from Saturday's humidity.

```{r message=FALSE, warning=FALSE}

# Visualizing Based-R

lm.weather <- lm(Temp_X ~ Humidity_X, data = weatherData)

plot ( Temp_X ~ Humidity_X
     , data = weatherData
     , xlab = "Humidity"
     , ylab = "Temperature"
     , main = "Temperature vs Humidity"
     , col  = "skyblue"
     , pch  = 16
     , cex  = 3 ) # dot size

abline(  lm.weather
       , col = "skyblue"
       , pch = 1
       , lwd = 3   # line width
       , lty = 2 ) # line type

points(  16.3
       , 82.33 # Humidity for Sep 18th (two days prior)
       , col = "firebrick"
       , pch =16
       , cex = 3 ) # dot size

text(  17
     , 85
     , "  Predicted Temperature \n (88.8 Fahrenheit)"
     , pos=3
     , cex=1.0
     , col="firebrick") 

```


<br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br />

## Data

### Table and Analysis

```{r message=FALSE, warning=FALSE}

datatable( weatherData
         ,extensions="Responsive"
         ,options=list(lengthMenu=c(10,5,10))
          ) 

```

Hypothesis:


$$
\left.\begin{array}{ll}
H_0: \beta_1 = 0 \\  
H_a: \beta_1 \neq 0
\end{array}
\right\} \ \text{Slope Hypotheses}
$$

$$
\left.\begin{array}{ll}
H_0: \beta_0 = 0 \\  
H_a: \beta_0 \neq 0 \\
\alpha =< 0.05
\end{array}
\right\} \ \text{Intercept Hypotheses}
$$

From the data collected, we get this estimated line representing the true regression model assumed by a regression analysis.

$$
\underbrace{\hat{Y_i}}_{Temperature}
=
\underbrace{\beta_0}_{Y-Intercept}
+
\underbrace{\beta_1}_{Slope}
\overbrace{X_i}^\text{Humidity}
$$
```{r message=FALSE, warning=FALSE}

pander(summary(lm.weather))

```

After computing the linear model, we were able to estimate the Y-intercept and slope for the equation, along with other meaningful values.

The y-intercept is __88.83__.
This means that when the percentage of humidity is zero, we estimate the highest temperature to be 88.83 degrees.

The slope has a value of __-0.4__.
This means that For every one whole drop in humidity, the temperature increases 16.34 grades.
Both variables have an indirectly proportional relation.

The R-squared value is __0.092.5__.
This is a really good value for our regression. This value represent the explanation capacity from the the linear model. This means that we have about 92.5% accuracy in our p-value.

The P-value has a value of __0.043__.
The P-Values is significant. This means that we reject the null hypothesis, accepting our hypothesis. 

From the regression model summary, we obtain the values of the regression line. The equation is written as follow:

<center>$\hat{Y_i}=88.83-0.4$ $X_i$</center>

<br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br />

## Assumptions

### Assumptions

```{r message=FALSE, warning=FALSE}

par(mfrow=c(1,3))
plot(  lm.weather
     , which = 1)
qqPlot(  lm.weather$residuals
       , id = FALSE
       , main = "Q-Q Plot")
plot(  lm.weather$residuals
     , main = "Residuals vs Order")

```

* Residuals vs Fitted: This graph has no discernible patterns. The fitted-values are the points in the regression line predicted thought the linear model. The residual-values are the distance from the estimated-value and the actual-value, mathematically below. The relationship of these two values is not linearity in the graph, therefore this assumption is fulfilled.

<br /><br />

<center>$Residual_i = Y_i - \hat{Y}_1$ </center>

<br /><br />

* Q-Q Plot: The residuals appear to be normal and within one standard deviation away from the regression line. Therefore this assumption is fulfilled.

* Residuals vs Order: This plot measures each residual and their order of occurrence. We are looking for no define trends in the plot. Since the plot show values with a slight down right-hand side trend, __this assumption is not fulfilled__.

Now that we have analyzed the 3 assumptions, and fail to approve one of them, we can not rely on the results from the Prediction

<br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br />

## Prediction

### Prediction

In the regression line, the **Estimated Temperature** was used as the responsive variable, and the **Humidity** is used as the explanatory variable.

In order to calculate the estimated maximum Temperature, the explanatory variable used is the pre-calculated **Humidity** value given to us in the data base. That **Humidity** is **16.3%**.

The mathematical equation for the estimated slope goes as follows:

$\underbrace{\hat{Y}_i}_\text{Temperature} = 88.83 - -0.4$ $\underbrace{X_i}_\text{Humidity}$

```{r message=FALSE, warning=FALSE}

pander(predict.lm(lm.weather, data.frame(Humidity_X=16.3)))

# lm.weather <- lm(Temp_X ~ Humidity_X, data = weatherData)
# predict.lm(lm.weather, data.frame(Humidity_X=16.3), interval="prediction")

```

After running a prediction model, the Temperature for two days from Monday September 20th, 2021, is __82.33__.

<!--

###Prediction update with transformation curve.

```{r message=FALSE, warning=FALSE}

# wrangling
plot(Temp_X ~ Humidity_X
     , data = weatherData
     , xlab = "Humidity"
     , ylab = "Temperature"
     , main = "Maximum Temperature vs Humidity"
     , col="blue"
     , pch=16)

lm.weather <- lm(Temp_X ~ Humidity_X
                 , data = weatherData)
b <- coef(lm.weather)
curve(  b[1] + b[2]*x
      , add=TRUE
      , col="hotpink")

pred.lm.weather <- predict(lm.weather                            , data.frame(Humidity_X = .53)
          , interval = "prediction")
abline(  h   = pred.lm.weather
       , lty = 2
       , col = "hotpink")
abline(  v=0.53
       , lty=2
       , col="hotpink")
# Fancier way to get predictions on the graph

lines(  c(0.53,0.53)
      , pred.lm.weather[2:3]
      , lwd = 8
      , col = rgb(1,0.41,.706,0.3))


```

-->

<br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br />

## Source

### Source

Data source: [https://www.timeanddate.com/weather/@5605242/historic?month=1&year=2020](https://www.timeanddate.com/weather/@5605242/historic?month=1&year=2020)

Time and Date AS is based just outside Stavanger, Norway, where the team work together to operate [timeanddate.com](https://www.timeanddate.com/), the world's top-ranking website for time and time zones. With free online services.

The German and Norwegian sites, [timeanddate.de](https://www.timeanddate.de/) and [timeanddate.no](https://www.timeanddate.no/), offer the same high level of accuracy and reliability as the English language site.



```{r message=FALSE, warning=FALSE}
# t.lm <- lm(  data = airquality
#            , Temp ~ Wind)
# summary(t.lm)
# plot(  Temp ~ Wind
#      , data= airquality
#      , col = "ivory3"
#      , pch = 18)
# abline(t.lm)

#
```


