---
title: "Chevrolet Corvette vs. Cadillac CTS-V"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

## Overview

The Chevrolet Corvette and the Cadillac CTS-V are two very well known and highly demanded cars today.
In this analysis, we can appreciate the comparison in their price over time. The analysis includes the initial price when new and the price throughout time.
It make assumptions such as the aggressiveness of the driving, and car load demand are equal for both vehicles.

```{r, include=FALSE}
# load packages
library(tidyverse)
library(pander)
library(DT)
library(dplyr)
# load dataset
CarPrices <- read.csv("../../Data/CarPrices.csv", header=TRUE)
```

In this graph, we can see the slight difference in slopes. We can also appreciate that Chevrolet's price drops faster than Cadillac's price.

```{r message=FALSE, warning=FALSE}
# data base filtered
modeltype <- CarPrices %>% 
  filter(
    Make == 'Chevrolet' | Make == 'Cadillac',
    Model == 'Corvette'    | Model == 'CST-V',
    Trim == 'Coupe 2D'  | Trim == 'Sedan 4D')

lm.2lines <-lm(Mileage ~ Price + Make + Price:Make, data=modeltype)

b <- coef(lm.2lines)

# graph
ggplot(modeltype, aes(y=Mileage, x=Price, color=factor(Make))) +
  geom_point(pch=21, bg="gray83") +
  stat_function(fun = function(x) b[1] + b[2]*x, color="skyblue") +
  stat_function(fun = function(x) (b[1]+b[3]) + (b[2]+b[4])*x,color="orange") +
  scale_color_manual(name="Transmission (am)", values=c("skyblue","orange")) +
  labs(title="TComparisson Model between Chevrolet Corvette and Cadillac CST-V ") 







```

## Model

The model description uses a quantitative X1i variable which is going to be the mileage, and a 0,1 indicator variable X2i. In this case, the indicator variable X2i, is either Chevrolet or Cadillac, producing two separate lines for each. The expression is written as follow:


$$
  \underbrace{Y_i}_{\text{Price}} = \overbrace{\beta_0 + \beta_1 \underbrace{X_{i1}}_{\text{Mileage}}}^{\text{Chevrolet Line}} + \overbrace{\beta_2 \underbrace{X_{i2}}_{\text{Cadillac}} + \beta_3 \underbrace{X_{i1} X_{i2}}_{\text{Interaction}}}^{\text{Cadillac Line}} + \epsilon_i
$$


where $\epsilon_i\sim N(0,\sigma^2)$ and $X_{i2} = 0$ when the vehicle is a Chevrolet and $X_{i2} = 1$ when the vehicle is a Cadillac This forced 0, 1 encoding for $X_{i2}$ produces the following models.

<center>

| Vehicle | Value of $X_{i2}$ | Resulting Model   |
|---------|-------------------|-------------------|
| Chevrolet   | $X_{i2} = 0$      | $Y_i = \beta_0 + \beta_1 X_{i1} + \epsilon_i$ |
| Cadillac | $X_{i2} = 1$      | $Y_i = (\beta_0 + \beta_2) + (\beta_1 + \beta_3) X_{i1} + \epsilon_i$ |

</center>

In this model, $\beta_2$ is the y-intercept difference and $\beta_3$ is the difference i slops for each vehicle.

## Hypothesis: Check y-Intercept Equality


If in the combined regression model $\beta_2$ is zero, then the average cost of the Chevrolet and Cadillac (y-intercepts) are the same for the same.
If in the combined regression model $\beta_2$ is more than zero, then a brand new Cadillac costs more than a brand new Chevrolet.
If in the combined regression model $\beta_2$ is smaller than zero, then the Cadillac costs less than the Chevrolet.

$$
  H_0: \beta_2 = 0 \quad \text{(same brand new price average)} \\
  H_a: \beta_2 \neq 0 \quad \text{(different brand new price average)}
$$

## Hypothesis: Check Slopes Equality

$$
  H_0: \beta_3 = 0 \quad \text{(same depreciation price)} \\
  H_a: \beta_3 \neq 0 \quad \text{(different depreciation price)}
$$

For the Y-Intercept and Slope hypothesis the significance level will be $\alpha=0.05$.

## Analysis

Notice that the graph interactions between the price vs mileage can be rectify with the Regression Summary Table
In the linear regression analysis, each set of points gets its to have its own slope. this means that the slope of the lines are different for Chevrolet and Cadillac.
In the model, the slope for Chevrolet is 0.1617, while the slop for Cadillac is 0.1585. Since Chevrolet has a higher slop, the price drop is more drastic than the Cadillac vehicle.

```{r message=FALSE, warning=FALSE}
pander(lm.2lines, caption="Regression output for Price on Mileage according to Model Type")

```



## Conlutions

Based on the analysis done, we can conclude that on average, the Cadillac CST-V Sedan 4D Sedan has an initial price more expensive than the Chevrolet Corvette Coupe 2D Coupe. We can also conclude that on average, the rate of price depreciation for the Chevrolet Corvette Coupe 2D Coupe is higher than the Cadillac.

```{r message=FALSE, warning=FALSE}

```


## Assumptions

The assumptions for the analysis are not meet. This means that we cannot make conclusions based on the statistical analysis demonstrated above.
On our first graph 'Residuals vs Fitted' we do not have a straight regression line, therefore we fail to check that assumption.On the 'Normal Q-Q' plot, we see some outliers and dots of data outside the graphical line, therefore we fail to check this assumption as well.

```{r message=FALSE, warning=FALSE}
par(mfrow=c(1,3))
plot(lm.2lines,which=1:2)
plot(lm.2lines$residuals)

mtext(side=3,text="Q-Q Plot of Residuals")
plot(lm.2lines$residuals, type="b")
mtext(side=3, text="Residuals vs. Order")
```


## Data Source

Data used to make the analys.

```{r message=FALSE, warning=FALSE}
modeltype <- CarPrices %>% 
  filter(
    Make == 'Chevrolet' | Make == 'Cadillac',
    Model == 'Corvette' | Model == 'CST-V',
    Trim == 'Coupe 2D'  | Trim == 'Sedan 4D' 
  )
datatable(modeltype)
```





















