---
title: "Simple Linear Regression"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
# libraries
library(car)
library(DT)
library(tidyverse)
library(pander)
library(mosaic)

# load data

dust <- read.csv("../../Data/Dust_Measurement.csv", header=TRUE)

```


<!-- Directions

Come up with a simple linear regression of your own creation. 

* Write the Background and clearly state your question. Then collect data and perform the linear regression analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.

* You should collect both the quantitative response variable and the quantitative explanatory variable yourself. This can be data you collect from 
    - the internet (like weather data, Facebook data, or sports data) 
    - your own experiment 
    - your own observational study
* However,     
    - The data should not be a single pre-existing dataset. It must be something that you piece together yourself.

* You need at least $n=3$ $(X,Y)$ pairs to perform the regression. However, your regression will only be useful if you can get around $n=10$ or more observations.

-->


## Background

In this analysis I am analyzing the thickness and weight for a specific area. I took a many samples of dust from the grown and used it to compare their weight before and after. I recorded the initial weight value, and then I used a filter machine that removes any big chunks of dust and then I recorded the end weight of the left over sand of that sample dust.

I am using a Simple Linear Regression of two numeric variables to see if there is a tendency in the data, as well as to see if there is a weights relation between the original sample and the particles after the filter.

The initial weight was recorded as the X-axis, and the measurement of the filter weight was measured in the Y-axis. The data collected comes from measuring the weight of dust in a 1 by 1 meters of area.


$$
H_0: \beta_1 = 0 \\
H_a: \beta_1 \neq 0\\
\alpha = 0.05
$$

The hypothesis can also be read as follows:

$$
  H_0: \text{The initial weight does not have a relation on the filter weight}
$$ 
$$
  H_a: \text{The intial weight has a relation on the filter weight}
$$

This analysis attempts to model the actual weight of individuals according to their reported weight using a linear regression. The model assumed by the regression analysis was

$$
  \underbrace{Y_i}_\text{Actual Weight} = \beta_0 + \beta_1 \underbrace{X_i}_\text{Reported Weight} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2) 
$$


## {.tabset .tabset-pills .tabset-fade}

### Hide Data

### Show Data {.tabset}

```{r message=FALSE, warning=FALSE}
datatable(dust)
```

## Analysis

The scatterplot is showing the relationship between the initial weight and filtered weights. It shows a light positive trend. The dots are dispersed around the line, but connected enough to make a positive trend.

```{r message=FALSE, warning=FALSE}

dust_regression <- lm(Y ~ X, data = dust)

ggplot(dust_regression, aes(x = X, y = Y)) +
  geom_point(size = 3.5, color = "orange", alpha = 0.5) +
  geom_smooth(
    method = "lm",
    formula = y~x,
    se = FALSE,
    color = "green",
    size = 1) +
  theme_bw() +
  labs(
    x = "Initial Weight",
    y = "Filtered Weight") +
  ggtitle("Change in Dust Weight")
```

This shows that (on average) for every additional kg of dust, the filter dust increase an additional 0.4 (3.899) kg.

In the Regression calculation, we come to know that the estimated slope of this regression is 0.3899, which is significantly different from zero (p=0.007013). Since the slope aligns with the P-Value, we can assume that we have sufficient evidence to reject the null hypothesis. the A complete table showing the summary of the full regression results are as follows:

```{r message=FALSE, warning=FALSE}

pander(summary(dust_regression))

```

These graph are checking assumptions. The "Residuals vs Fitted" graph does not have a linear relationship nor a constant variance, so we fail to make this assumption. Along with the "Residuals vs Fitted" graph, the "Normal Q-Q" visual plot is also not consider a normal probability plot, therefore the error terms are also assumed to violate the normality assumption. The "Residuals versus Order Plot" is not showing a trend, which allows to validate this assumption.

We fail to validate two out of the three assumptions based on the graphs. 

```{r message=FALSE, warning=FALSE}

par(mfrow=c(1,3))
plot(dust_regression,which=1:2)
plot(dust_regression$residuals)

```


## Interpretation

Based on the analysis, the p-value is smaller than $alpha$, indicating that we reject the null hypothesis.
We can assume that the weights of dust have a positive relation with their weights filtered. The more heave the dust, the more heavy the filter dust is as well.

Due to the assumptions not being meet, we cannot affirm that this trend is accurate. Therefore we cannot assume that the initial weight of dust has a correlation with the filter weight of dust for each sample.




